---
title: "Test of garden hunting hypothesis for mammals in La Gran Sabana, Venezuela using occupancy models"
author: "Stachowicz, I; Ferrer-Paris, J.R.; Sanchez-Mercado, A. (in prep)"
date: December 2, 2020
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE)
require(unmarked)
require(AICcmodavg)
require(chron)
require(raster)
require(cluster)
require(MuMIn)
require(dplyr)
require(ggplot2)


if (Sys.getenv("WORKDIR") == "") {
   ## for Izza:
   work.dir <- "D:/PROJECTS/Gran Sabana/Metodologia"
   script.dir <- "???"
} else {
   ## for JR:
   work.dir <- Sys.getenv("WORKDIR")
   script.dir <- Sys.getenv("SCRIPTDIR")
}

setwd(work.dir)
GIS.data <- sprintf("%s/Rdata/GIS.rda",script.dir)
load(GIS.data)

```

# Methods

## Model definition

## Covariates of probability of detection

dras: distance to animal tracks and trails

sfrz: sampling effort (nr. of days camera was active)

date: date of sampling

## Covariates of probability of occupancy - frecuency of use

buf.fragmen: 1 km buffer of forest cover derived from LandSat time series (Hansen et al. 2013)

dcon: distance to nearest conuco


```{r}
ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(x=dcon,y=wcon,colour=as.factor(caza.celda>0))) + geom_point() +
labs(title="Distance to nearest conucos vs. density of conucos") +
labs(x="Distance (m)", y="Number of conucos (Inverse Distance Weighted)",colour="Hunting")
##ggsave("Fig-distnce-density-conucos.png")

```

```{r,eval=FALSE,echo=FALSE}
ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(x=dcon,fill=as.factor(caza.celda))) + geom_histogram(binwidth=500, alpha = .8,col='black') +
labs(title="Histogram for distance to conucos") +
labs(x="Distance (m)", y="Count")

ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(x=wcon,fill=as.factor(caza.celda))) + geom_histogram(alpha = .8,col='black') +
labs(title="Histogram for distance to conucos") +
labs(x="Distance (m)", y="Count")
```

```{r}
subset(camaras,bloque %in% sprintf("B%02s",1:6) & dcon>7000)
```

```{r,eval=FALSE,echo=FALSE}
 ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(x=buf.fragmen,fill=as.factor(caza.celda))) + geom_histogram(binwidth=5, alpha = .8,col='black') +
labs(title="Histogram for forest cover") +
labs(x="Percentage cover", y="Count")
```

```{r,eval=FALSE,echo=FALSE}
ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(x=buf.fragmen,fill=bloque)) + geom_histogram(binwidth=5, alpha = .8,col='black') +
labs(title="Histogram for forest cover") +
labs(x="Percentage cover", y="Count")
```



```{r}

ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(y=dcon,x=buf.fragmen,colour=as.factor(caza.celda>0))) +
geom_point() +
labs(title="Forest cover vs distance to conuco") +
labs(y="Distance (m)", x="Forest cover (%)",colour="Hunting")
##ggsave("Fig-bosque-caceria.pdf")


```



```{r}
ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(y=wcon,x=buf.fragmen,colour=as.factor(caza.celda),size=caza.celda)) +
geom_point() +
labs(title="Forest cover vs density of conuco") +
labs(y="Density", x="Forest cover (%)")



```

Eliminar tres puntos mas remotos (> 5km del conuco)


```r
ggplot(data=subset(camaras,bloque %in% sprintf("B%02s",1:6)), aes(y=idw.conucos,x=min.conucos,colour=as.factor(caza.celda),size=caza.celda)) +
geom_point()
```


# Results

## Goodness of fit

MacKenzie and Bailey Goodness-of-fit Test for  Royle-Nichols Occupancy Models.

```{r }
if (!exists("tbl1")) {
  tbl1 <- data.frame()
  for (arch in dir(sprintf("%s/Rdata/occuRN/",script.dir))) {
    load(sprintf("%s/Rdata/occuRN/%s",script.dir,arch))
  mi.spp <- sub(".rda$","",arch)
  UMF <- get(sprintf("UMF.%s",mi.spp))
  ts03 <- get(sprintf("ts03.%s",mi.spp))
  fm01 <- get(sprintf("fm01.%s",mi.spp))
  fm03 <- get(sprintf("fm03.%s",mi.spp))

   tbl1 <- rbind(tbl1,data.frame(spp=mi.spp,
   n.detect= sum(UMF@y,na.rm=T),
  ts03[c("chi.square","p.value","c.hat.est")],
  large.coefs=min(max(abs(coef(fm01))),max(abs(coef(fm03)))),
  large.SE=min(max(SE(fm01)),max(SE(fm03)))))

  }
  tbl1 %>% mutate(overdispersion=c.hat.est > 1 & c.hat.est <4, coef.problems=large.coefs>4 | large.SE>4) %>%
    mutate(lack.of.fit=p.value<.10 | p.value > .9 | c.hat.est > 4 | c.hat.est <0.30 ) -> tbl1

}

```

Symptoms of lack of fit for most species: Six models with extreme p-value or c-hat values, 19 models with large coeficients or standard errors.

```{r }
table(tbl1$lack.of.fit,tbl1$coef.problems)

```

These species show one or more signs of lack of fit, probably due to the low number of detections:

```{r }
tbl1 %>% filter(lack.of.fit) %>% select(1:7)

```

These species appear to have  a good fit but might have problems with large coeficients and standard errors
```{r }
tbl1 %>% filter(!lack.of.fit & coef.problems) %>% arrange(c.hat.est) %>% select(1:7)
```

For this species, the over-dispersion might be accounted for by using quasi-AICc
```{r }
tbl1 %>% filter(!lack.of.fit & !coef.problems & overdispersion) %>% arrange(c.hat.est) %>% select(1:7)
```

These species seem to have  a good fit and no signs of over-dispersion:

```{r }
tbl1 %>% filter(!lack.of.fit & !coef.problems & !overdispersion) %>% arrange(c.hat.est) %>% select(1:7)

```


## Model averaging

### Variable importance

```{r }
slc <- c('C.paca','C.alector', 'D.leporina', 'M.gouazoubira', 'D.kappleri', 'D.novemcinctus', 'L.rufaxilla', 'L.pardalis','C.thous','P.onca')
tst <- c('D.imperfecta','T.terrestris','T.major','M.americana')

ccs <- data.frame()
for (spp in c(slc,tst)) {
  print(spp)
  if (spp %in% c('C.alector','L.rufaxilla','T.tetradactyla')) {
    print(sw(get(sprintf("oms03.%s",spp))))
    mavg <- get(sprintf("mavg03.%s",spp))
  } else {
    print(sw(get(sprintf("oms01.%s",spp))))
    mavg <- get(sprintf("mavg01.%s",spp))
  }
  ccs <- rbind(ccs,data.frame(spp, coef(mavg,full=F),confint(mavg,full=F)))
}

```


```{r}
ccs  %>% filter(grepl('dcon',rownames(ccs))) %>% dplyr::arrange(coef.mavg..full...F.) -> ss
par(mar=c(4,8,3,1))
plot(ss$coef.mavg.,1:nrow(ss),xlim=c(-1.8,1.8),pch=19,xlab=expression(hat(beta)),ylab='',axes=F,main='distance to nearest conuco')
segments(ss$X2.5..,   1:nrow(ss),   ss$X97.5..,1:nrow(ss))
axis(1)
axis(2,1:nrow(ss),ss$spp,las=2,font=3)
box()
abline(v=0,lty=2,lwd=2,col=2)
# dev.copy(png,file='Fig-coefficient-distance.png')
# dev.off()
```


```{r,eval=FALSE,echo=FALSE}
ccs <- data.frame()
for (spp in c('D.leporina','L.pardalis','C.alector','C.thous','P.onca','D.novemcinctus','M.gouazoubira','D.kappleri','C.paca','L.rufaxilla')) {
  print(spp)
  if (spp %in% c('L.pardalis','D.novemcinctus','L.rufaxilla')) {
    print(sw(get(sprintf("oms13.%s",spp))))
    mavg <- model.avg(get(sprintf("oms13.%s",spp)), subset = delta < 10)
  } else {
    print(sw(get(sprintf("oms11.%s",spp))))
    mavg <- model.avg(get(sprintf("oms11.%s",spp)), subset = delta < 10)
  }
  ccs <- rbind(ccs,data.frame(spp, coef(mavg,full=F),confint(mavg,full=F)))
}

```

```{r,echo=FALSE,eval=FALSE}
ccs  %>% filter(grepl('buf.fragmen',rownames(ccs))) %>% dplyr::arrange(coef.mavg..full...F.) -> ss

plot(ss$coef.mavg.,1:nrow(ss),xlim=c(-2.3,2.3),pch=19,xlab='',ylab='',axes=F,main='buf.fragmen')
segments(ss$X2.5..,   1:nrow(ss),   ss$X97.5..,1:nrow(ss))
axis(1)
axis(2,1:nrow(ss),ss$spp,las=2)
box()
abline(v=0,lty=2,lwd=2,col=2)
```




## Hunting


```{r}



mtz <- data.frame()

Hv <- c('C.paca'=6.336,'C.alector'=4.630, 'D.leporina'=2.681, 'T.terrestris'=2.681,'T.major'=1.949, 'M.gouazoubira'=0.731,'M.americana'=0.731, 'D.kappleri'=0.244, 'D.novemcinctus'=0.244)
# reported hunted
slc <- c('C.paca','C.alector', 'D.leporina', 'T.terrestris','T.major', 'M.gouazoubira','M.americana', 'D.kappleri', 'D.novemcinctus')
# not reported hunted
tst <- c( 'L.rufaxilla','D.imperfecta', 'L.pardalis','C.thous','P.onca')


exc <- c('M.tridactyla','T.tetradactyla','P.maximus','P.jacquacu', # large coeficients / unrealistic predictions
  'N.nasua','D.marsupialis','T.pecari','C.olivaceus','L.wiedii',
  'P.concolor','E.barbara',
  "C.unicinctus", "H.hydrochaeris","O.virginianus","P.tajacu"  )  # excluded due to lack of fit
# levels(eventos$species)[!levels(eventos$species) %in% c(slc,tst,exc)]

for (k in c(slc)) {
  if (spp %in% c('C.alector','L.rufaxilla','T.tetradactyla')) {
      mtz <- rbind(mtz,data.frame(species=k,abundance=predict(get(sprintf("mavg03.%s",k)),type='state')$fit,caza=UMF@siteCovs$caza.celda))
    } else {
      mtz <- rbind(mtz,data.frame(species=k,abundance=predict(get(sprintf("mavg01.%s",k)),type='state')$fit,caza=UMF@siteCovs$caza.celda))
    }
}

mtz$hunting <- ifelse(mtz$caza==1,'yes','no')
mtz$hunting <- ifelse(mtz$caza>0,'yes','no')

library(grid)
text_Hv1 <- textGrob(sprintf("(%s)",Hv[1]), gp=gpar(fontsize=7))
text_Hv2 <- textGrob(sprintf("(%s)",Hv[2]), gp=gpar(fontsize=7))
text_Hv3 <- textGrob(sprintf("(%s)",Hv[3]), gp=gpar(fontsize=7))
text_Hv4 <- textGrob(sprintf("(%s)",Hv[4]), gp=gpar(fontsize=7))
text_Hv5 <- textGrob(sprintf("(%s)",Hv[5]), gp=gpar(fontsize=7))
text_Hv6 <- textGrob(sprintf("(%s)",Hv[6]), gp=gpar(fontsize=7))
text_Hv8 <- textGrob(sprintf("(%s)",Hv[8]), gp=gpar(fontsize=7))

# grouped boxplot
P <- ggplot(mtz %>% filter(), aes(x=species, y=abundance, fill=hunting)) +
    geom_boxplot(notch=F) + # or notch=T
    labs(title="Model prediction of abundance at sites with and without hunting") +
    labs(y=expression( hat(lambda)), x="",caption="Species (Hv values)") +
    theme(axis.text.x = element_text( size = 7, hjust = .5, vjust=.5, face = "italic"),
    plot.margin = unit(c(1,1,2,1), "lines")) +
    coord_cartesian(clip="off")
  P +  annotation_custom(text_Hv1,xmin=1,xmax=1,ymin=-0.5,ymax=-0.5) +
  annotation_custom(text_Hv2,xmin=2,xmax=2,ymin=-0.5,ymax=-0.5) +
   annotation_custom(text_Hv3,xmin=3,xmax=4,ymin=-0.5,ymax=-0.5) +
    annotation_custom(text_Hv5,xmin=5,xmax=5,ymin=-0.5,ymax=-0.5) +
    annotation_custom(text_Hv6,xmin=6,xmax=7,ymin=-0.5,ymax=-0.5) +
    annotation_custom(text_Hv8,xmin=8,xmax=9,ymin=-0.5,ymax=-0.5)


# vjust = c(.3,.3,.3,.7,.3,.7,.3,.3,.3,.3)
## ggsave("Fig-abundance-hunting.png",width=8,height=5)
# ggsave("Fig-abundance-hunting-with-notches.png",width=8,height=5)

```


```{r}



mtz <- data.frame()

for (k in c(tst)) {
  if (spp %in% c('C.alector','L.rufaxilla','T.tetradactyla')) {
      mtz <- rbind(mtz,data.frame(species=k,abundance=predict(get(sprintf("mavg03.%s",k)),type='state')$fit,caza=UMF@siteCovs$caza.celda))
    } else {
      mtz <- rbind(mtz,data.frame(species=k,abundance=predict(get(sprintf("mavg01.%s",k)),type='state')$fit,caza=UMF@siteCovs$caza.celda))
    }
}

mtz$hunting <- ifelse(mtz$caza==1,'yes','no')
mtz$hunting <- ifelse(mtz$caza>0,'yes','no')


# grouped boxplot
ggplot(mtz %>% filter(), aes(x=species, y=abundance, fill=hunting)) +
    geom_boxplot(notch=F) + # or notch=T
    labs(title="Model prediction of abundance at sites with and without hunting") +
    labs(y=expression( hat(lambda)), x="") +
theme(axis.text.x = element_text( size = 7, hjust = .5, vjust=.5, face = "italic"))

# ggsave("Fig-abundance-hunting-not-reported.png",width=8,height=5)

```

## results per species



### C. paca


Sum of AICc weights indicate a clear effect of p(sfrz) p(dras) and lam(dcon)

```{r }
summary(mavg01.C.paca)

re <- ranef(fm01.C.paca,K=30)

bup(re, stat="mean")           # Posterior mean
     bup(re, stat="mode")           # Posterior mode
     confint(re, level=0.9) # 90% CI
 plot(bup(re, stat="mode")      ,predict(fm01.C.paca,type='state')$Predicted)
colSums(confint(re, level=0.9) )
colSums(predict(fm01.C.paca,type='state'))



aggregate(cbind(bup(re, stat="mean"),confint(re, level=0.9)),list(y),sum )

aggregate(predict(fm01.C.paca,type='state'),list(y),sum )
```

Significant conditional coeficients for those parameters. Negative relationship with distance to conuco ("atracted") of conucos

### C. alector

Most support for p(dras)+p(sfrz)+lam(evi.mu)+lam(wcon), significant conditional coeficients for those parameters. Strong negative significative effect of conucos.

```{r }
summary(mavg03.C.alector)

```


```{r }
prd <- predict(mavg03.C.alector,type='state')
##prd <- ranef(mavg01.C.alector,type='state')
dtf <- data.frame(prd=prd$fit,se=prd$se.fit,caza=factor(UMF@siteCovs$caza.celda>0),wcon=UMF@siteCovs$dcon,bsq=UMF@siteCovs$buf.fragmen)


ggplot(dtf, aes(x=wcon, y=prd,colour=factor(caza))) +
  geom_errorbar(aes(ymin=prd-se, ymax=prd+se), colour="grey", width=0.025) +
    geom_point() + xlab("distance to conucos") + ylab("Predicted values") + labs(colour='Hunting')

    ggplot(dtf, aes(x=bsq, y=prd,colour=factor(caza))) +
      geom_errorbar(aes(ymin=prd-se, ymax=prd+se), colour="grey", width=0.025) +
        geom_point() + xlab("bosque") + ylab("Predicted values") + labs(colour='Hunting')

```
`

### D. leporina
Used linear model for EVI. Most support for p(sfrz)+lam(evi.mu), significant conditional coefficients for those parameters. Weak negative ("avoids") non-significant effect of conucos

```{r }
summary(mavg01.D.leporina)


```

```{r }
prd <- predict(mavg01.D.leporina,type='state')

dtf <- data.frame(prd=prd$fit,se=prd$se.fit,caza=factor(UMF@siteCovs$caza.celda>0),wcon=UMF@siteCovs$dcon,bsq=UMF@siteCovs$buf.fragmen)


ggplot(dtf, aes(x=wcon, y=prd,colour=factor(caza))) +
  geom_errorbar(aes(ymin=prd-se, ymax=prd+se), colour="grey", width=0.025) +
    geom_point() + xlab("distance to conucos") + ylab("Predicted values") + labs(colour='Hunting')


```



### L. rufaxilla

Sum of AICc weights indicate a clear effect of p(sfrz) and large support for lam(buf.fragmen). Almost half of the models support lam(dcon)

```{r }
summary(mavg03.L.rufaxilla)

```

Significant conditional coeficients for p(sfrz). Negative relationship with distance to conuco ("atracted") but non-significant effect of conucos



```{r }
prd <- predict(mavg03.L.rufaxilla,type='state')

```
### C.thous

Used linear model for EVI. Most support for p(sfrz)+lam(evi.mu), significant conditional coefficients for lam(evi.mu) (negative association). Weak negative ("avoids") non-significant effect of conucos

```{r }
summary(mavg01.C.thous)

```




### L. pardalis

Most support for null model.

```{r }
summary(mavg01.L.pardalis)

```



### D. kappleri

Most support for p(sfrz)+lam(evi.mu) but no significant conditional coefficients . Very weak  non-significant effect of conucos.


```{r }
summary(mavg01.D.kappleri)

```

### D. novemcinctus

```{r }
summary(mavg01.D.novemcinctus)

```


### P.onca


```{r }
summary(mavg01.P.onca)

```

### M.gouazoubira


```{r }
summary(mavg01.M.gouazoubira)

```


### TO BE TESTED


```{r }
summary(mavg01.D.imperfecta)
summary(mavg01.T.terrestris)
summary(mavg01.M.tridactyla)
summary(mavg03.T.tetradactyla)
```




```{r }
prd <- predict(mavg01.C.thous,type='state')
dtf <- data.frame(prd=prd$fit,se=prd$se.fit,caza=UMF@siteCovs$caza.celda,dcon=UMF@siteCovs$dcon,bsq=UMF@siteCovs$buf.fragmen)


ggplot(dtf, aes(x=dcon, y=prd,colour=factor(caza))) +
  geom_errorbar(aes(ymin=prd-se, ymax=prd+se), colour="grey", width=0.025) +
    geom_point() + xlab("distance to conucos") + ylab("Predicted values") + labs(colour='Hunting')

    ggplot(dtf, aes(x=bsq, y=prd,colour=factor(caza))) +
      geom_errorbar(aes(ymin=prd-se, ymax=prd+se), colour="grey", width=0.025) +
        geom_point() + xlab("forest cover") + ylab("Predicted values") + labs(colour='Hunting')

```










### M. americana


### T.tetradactyla






### T.terrestris

Very large standard errors
```{r }
summary(mavg03.T.terrestris)

```


```{r }
prd <- predict(mavg03.T.terrestris,type='state')
dtf <- data.frame(prd=prd$fit,se=prd$se.fit,caza=factor(UMF@siteCovs$caza.celda>0),dcon=UMF@siteCovs$dcon,bsq=UMF@siteCovs$buf.fragmen)
ggplot(dtf, aes(y=dcon, x=bsq,size=prd,colour=factor(caza))) +
    geom_point() + ylab("distance to conucos  (standardised values)") + xlab("Forest cover (standardised values)") + labs(colour='Hunting',size='Predicted values')


```


### D.imperfecta



### M.gouazoubira


### N.nasua


### M.tridactyla



### P. onca


### E.barbara

Large coefficients lead to large, unrealistic abundance predictions

```{r }
summary(mavg01.E.barbara)

```
